Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-08-14 Multiresolution_framework/ltx/response_to_reviews.tex

\section{Handling editor}
\begin{enumerate}
\item Whilst the paper does fit the scope of NI, I agree with R1 and R3, that explicit advantages of going to a multiresolution frame - in addition to the conceptual ones - compared to the prior work based on Gaussian basis functions needs clarification and elucidation.

\emph{\ken{Guys, I think the editor and reviewers 1 and 3 will be convinced by showing, just through an example, the added accuracy obtained by using the MRA over the Gaussian approximation. Is it possible to run such a comparison? Also, comparing with the results of our previous NI paper - can strengthen the argument that when operating in the linear region this approach can be superior.}}

\item  In addition, at least a ``proof of principle" inversion from empirical data would enhance the potential impact of the paper.

\dean{I think we should ignore this one. Otherwise, I have drafted a response.}
\emph{\ken{Guys - Not sure what the editor means by this - but I feel it would help to highlight further at the start of section 7 that the data is generated by another accepted model and not the model we are trying to estimate. Possibly - not using splines to obtain the mexican hat might make it even clearer.}}

The authors agree that the use of empirical data will strengthen the paper. Unfortunately, the authors do not have access to the appropriate data at present. However, we understand that several other groups are now collecting high spatiotemporal resolution data that is appropriate for the methods developed in this paper. We believe that the methods developed in this paper are equally important as the development of high spatiotemporal resolution recording techniques and together, these approaches have enormous potential to answer some of the biggest questions in neuroscience. The ability to use mathematical models as filters, or lenses into the brain if you like, is still a relatively new concept to many members in our community. This new lens allows us to image aspects of neurophysiology are normally hidden in electrical recording. Therefore, we believe it is important to continue improve our methods that will lead to the next era in systems neuroscience.

### 2012-08-14 Multiresolution_framework/ltx/response_to_reviews.tex

Parham Aram.

\section{Handling editor}
\begin{enumerate}
\item Whilst the paper does fit the scope of NI, I agree with R1 and R3, that explicit advantages of going to a multiresolution frame - in addition to the conceptual ones - compared to the prior work based on Gaussian basis functions needs clarification and elucidation.

\emph{\ken{Guys, I think the editor and reviewers 1 and 3 will be convinced by showing, just through an example, the added accuracy obtained by using the MRA over the Gaussian approximation. Is it possible to run such a comparison? Also, comparing with the results of our previous NI paper - can strengthen the argument that when operating in the linear region this approach can be superior.}}

\dean{Parham: Maybe you can make a list of key differences, highlighting the challenges, the technical details, and the benefits that the MRAIDE offers.}

@@

\item  In addition, at least a ``proof of principle" inversion from empirical data would enhance the potential impact of the paper.

\dean{I think we should ignore this one. Otherwise, I have drafted a response.}
\emph{\ken{Guys - Not sure what the editor means by this - but I feel it would help to highlight further at the start of section 7 that the data is generated by another accepted model and not the model we are trying to estimate. Possibly - not using splines to obtain the mexican hat might make it even clearer.}}

\emph{The authors agree that the use of empirical data will strengthen the paper. Unfortunately, the authors do not have access to the appropriate data at present. However, we understand that several other groups are now collecting high spatiotemporal resolution data that is appropriate for the methods developed in this paper. We believe that the methods developed in this paper are equally important as the development of high spatiotemporal resolution recording techniques and together, these approaches have enormous potential to answer some of the biggest questions in neuroscience. The ability to use mathematical models as filters, or lenses into the brain if you like, is still a relatively new concept to many members in our community. This new lens allows us to image aspects of neurophysiology are normally hidden in electrical recording. Therefore, we believe it is important to continue improve our methods that will lead to the next era in systems neuroscience.}

@@

        \item The authors say the paper at hand is an ``extension'' of Freestone et al. by which I understand they mean the use of a wavelet as opposed to a Gaussian basis for the decomposition.

	\emph{Response here.}

        \item Since the methodology used is the same as in that earlier Neuroimage paper, the paper at hand repeats several of  the results of that previous paper, see sections 3 and 4: equations (2)-(5) and (20) -(33) have already appeared or are very similar to equations in Freestone et al.

\dean{These first three things are the same point from Reviewer 1.}
	\emph{The earlier papers (Freestone et al., 2011, Dewar et al., 2009) both propose solutions to representing a spatiotemporal field with a finite number of basis functions. Together with this paper, we are building tools that deal with increasing levels of complexity in the spatiotemporal dynamics. The first paper describes key work in the PhD thesis of Dewar. The second in the PhD thesis of Freestone. This paper covers the some of the key results from the PhD thesis of Aram. Each of these steps was a major undertaking and should not be under estimated. There is similarities with the frameworks with all of these papers, where the field is represented by a basis function decomposition. The challenges in forming a multiresolution state-space model ... }

@@

        \item Section 5 is a review of known results and standard background material from wavelet theory and includes a simple calculation of the Fourier transform of the particular basis used here, which would be more appropriate for an appendix (cf appendix H of the Freestone et al. paper where the Fourier transform of the Gaussian basis functions is calculated).

		\emph{\parham{Section 5 summarizes important characteristics of B-Spline scaling and wavelet functions which are collected from different references, with an attempt to clarify the inner product and convolution of theses functions (please see R2 comment on ``B-spline calculus''). We think that this brief review can familiarize interested readers  with  B-spline scaling and wavelet functions and direct them to the important literature on this matter. As the reviewer suggested the calculation of the Fourier transform is now moved to the appendix.}}

\emph{\dean{We understand that this section will not be the central focus of the paper for the non-technical reader. This was the motivation for the comments in the section introduction, where we state... Although we appreciate R1 comments and suggestions, since R2 commented on the usefulness of this section we would very much like to leave it in. }}

### 2012-08-14 Multiresolution_framework/ltx/response_to_reviews.tex

        \item Since the methodology used is the same as in that earlier Neuroimage paper, the paper at hand repeats several of  the results of that previous paper, see sections 3 and 4: equations (2)-(5) and (20) -(33) have already appeared or are very similar to equations in Freestone et al.

\dean{These first three things are the same point from Reviewer 1.}
	\emph{The earlier papers (Freestone et al., 2011, Dewar et al., 2009) both propose solutions to representing a spatiotemporal field with a finite number of basis functions. Together with this paper, we are building tools that deal with increasing levels of complexity in the spatiotemporal dynamics. The first paper describes key work in the PhD thesis of Dewar. The second in the PhD thesis of Freestone. This paper covers the some of the key results from the PhD thesis of Aram. Each of these steps was a major undertaking and should not be under estimated. There is similarities with the frameworks with all of these papers, where the field is represented by a basis function decomposition. The challenges in forming a multiresolution state-space model ... }

        \item Section 5 is a review of known results and standard background material from wavelet theory and includes a simple calculation of the Fourier transform of the particular basis used here, which would be more appropriate for an appendix (cf appendix H of the Freestone et al. paper where the Fourier transform of the Gaussian basis functions is calculated).

		% \emph{\parham{Section 5 summarizes important characteristics of B-Spline scaling and wavelet functions which are collected from different references, with an attempt to clarify the inner product and convolution of theses functions (please see R2 comment on ``B-spline calculus''). We think that this brief review can familiarize interested readers  with  B-spline scaling and wavelet functions and direct them to the important literature on this matter. As the reviewer suggested the calculation of the Fourier transform is now moved to the appendix.}}

\emph{\dean{We understand that this section will not be the central focus of the paper for the non-technical reader. This was the motivation for the comments in the section introduction, where we state... Although we appreciate R1 comments and suggestions, since R2 commented on the usefulness of this section we would very much like to leave it in.} \parham{As the reviewer suggested the calculation of the Fourier transform is now moved to the appendix.}}

### 2012-08-14 Multiresolution_framework/ltx/response_to_reviews.tex

        \item Also, Appendix A is the analogue of Appendix D of Freestone et al. and Appendix C includes results known from the literature (as the authors point out).

\emph{\parham{In order to formulate the state-space representation of the multi-resolution neural field model the inner product of the B-spline functions are required. We believe that it is necessary to provide the interested reader with the required tools to implement the algorithm. Appendix D of Freestone et al. and Appendix A herein are only similar in the sense that the inner product of two basis functions are calculated. In Appendix D of Freestone et al. two  n-dimensional Gaussians are considered while here the inner product of two B-spline functions is calculated. As for Appendix C, although authors have no objections to remove the presented derivations, we think that it would be informative for the broad audience of NeuroImage to include the derivations of some of well-known formulations in the EM algorithm.} }

### 2012-08-14 Multiresolution_framework/ltx/manuscript.tex
correction to Table 1
drafting the choice of observation kernel

One snap shot of the field reconstruction, at different levels of approximation is given in Fig.~\ref{fig:FieldEstimates100}, showing small discrepancy between the actual and estimated field at $j=3$. From this figure, it can be observed that the overall trend of the underlying field can be also captured at the coarser levels.
\subsection{Experiment II}
To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The mean over time of the root mean square error (MRMSE) over space of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution finite element model representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. By decomposing both the spatial field and the connectivity kernel using a wavelet decomposition, it becomes possible to represent small scale details in the field at the same time as large scale details. This work is significant, as the ability to create patient-specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling, where the dynamics and connectivity structure differs at different spatial scales~\citep{Jirsa2002,Jirsa2004,Breakspear2005,Qubbaj2009,Schultze-Kraft2011}. 
\parham{\subsection{Model reduction}
Different basis functions can be used to reduce the neural field formulations for state and parameter estimation problem. An alternative is Gaussian basis functions  adopted by \cite{Freestone2011} to decompose the field and connectivity kernel. One clear advantage of B-spline scaling and wavelet functions is the MRA capability, allowing the simultaneous representation of the neural field at different spatial scales. B-spline scaling and wavelet functions are compactly supported, eliminating  numerical errors caused by the Gaussian basis function. The partition of unity of these basis functions enable the reconstruction of the neural field when it saturates, this may be beneficial to model neurological disorders such as epilepsy. Attaining fine scale approximation of the neural field using Gaussian basis functions results into an ill-posed problem as the number of states will be increased significantly, however, this is not the case when using B-spline scaling and wavelet functions due to their semi-orthogonality property.}

@@

		$\boldsymbol\theta$&Vector of connectivity kernel parameters composed of $\alpha_{j_0,l}$ and $\beta_{j,l}$& mV.spike$^{-1}$\\ 

>>>>

One snap shot of the field reconstruction, at different levels of approximation is given in Fig.~\ref{fig:FieldEstimates100}, showing small discrepancy between the actual and estimated field at $j=3$. From this figure, it can be observed that the overall trend of the underlying field can be also captured at the coarser levels.
\subsection{Experiment II}
To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The mean over time of the root mean square error (MRMSE) over space of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution finite element model representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. By decomposing both the spatial field and the connectivity kernel using a wavelet decomposition, it becomes possible to represent small scale details in the field at the same time as large scale details. This work is significant, as the ability to create patient-specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling, where the dynamics and connectivity structure differs at different spatial scales~\citep{Jirsa2002,Jirsa2004,Breakspear2005,Qubbaj2009,Schultze-Kraft2011}. 
\parham{\subsection{State-space representation}
Different basis functions can be used to reduce the neural field formulations for state and parameter estimation problem. An alternative is Gaussian basis functions  adopted by \cite{Freestone2011} to decompose the field and connectivity kernel. One clear advantage of B-spline scaling and wavelet functions is the MRA capability, allowing the simultaneous representation of the neural field at different spatial scales. B-spline scaling and wavelet functions are compactly supported, eliminating  numerical errors caused by the Gaussian basis function. The partition of unity of these basis functions enable the reconstruction of the neural field when it saturates, this may be beneficial to model neurological disorders such as epilepsy. Attaining fine scale approximation of the neural field using Gaussian basis functions results into an ill-posed problem as the number of states will be increased significantly, however, this is not the case when using B-spline scaling and wavelet functions due to their semi-orthogonality property. \cite{Freestone2011} showed experimentally that  Gaussian functions are suitable choice for modeling  observation kernel when intracranial measurements are considered. B-splines are also suitable for modeling observation kernel due to their close approximation to the Gaussian kernel. B-splines converge to the Gaussian function as the order of the B-spline $m$ tends to
infinity.}

@@

		$\boldsymbol\theta$&Vector of connectivity kernel parameters composed of $a_{j_0,l}$ and $c_{j,l}$& mV.spike$^{-1}$\\ 
